### 2016-11-29 dissertation/unsat.tex
dissertation v2:major chapters completed

@@

\subsection{Exploiting UNSAT cores for abstraction refinement}

UNSAT cores can find broad applications in circuit verification. Many abstraction refinement techniques require 
information mining from UNSAT proofs of intermediate abstractions. 
Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain that how an UNSAT proof is
utilized in such techniques.

% TODO explain BMC

@@

\begin{figure}[hbt]
\centering{
\includegraphics[width=5.5in]{newfig/refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
Assume that we are given a sequential circuit with $n$ latches as shown in Fig.\ref{fig:refine}(a). 
This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
$l_1,\dots,l_n$. Algorithm \ref{alg:absrefine} describes an approach to check if machine $M$ violates property $p$.
This algorithm relies on $k$-BMC technique, which works on the basis of CNF-SAT solving.
The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is $SAT$ or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
$SAT$, then a counterexample is found that violates property $p$. If the result is $UNSAT$, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fixpoint detected.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is $UNSAT$.
If there are some latch variables not included in this UNSAT core (denoted by $L_{abs}$), then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Fig.\ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

The second ``if-else" branch means: if we do an ordinary model checking on the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an over-approximation
of $M$, such that $(M'\implies p) \implies (M\implies p)$. If we find a violation on abstracted machine, then this
abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which will encounter the ``bit-blasting" problem
when the size of datapath (number of latches in Fig.\ref{fig:refine}) is very large. \textbf{Here we propose a totally new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation based on our experience:}

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While computing GB over finite fields is exponential in the number of variables, GB  computation 
is observed to be more efficient for UNSAT problems.} 
The reason is discussed as follows:
\begin{Theorem}
{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow reducedGB(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high
time complexity and is usually not practicable. One reason is that the size of GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is 1, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

\subsection{A Demonstration of Motivating Example}
One  important research topic about UNSAT problems is to identify UNSAT cores efficiently.
An UNSAT core in a CNF formula denotes a subset of clauses which is still unsatisfiable. Here
we re-define this concept in algebraic geometry:
\begin{Definition}
Assume a set of polynomials $F$ and its subset $F_s\subset F$. If ${\bf V}(\langle F\rangle) = {\bf V}
(\langle F_s\rangle) = \emptyset$, we call $F_s$ an {\bf UNSAT core} of $F$. Additionally, if
$F_s$ contains no other UNSAT core, we call it a {\bf minimal} UNSAT core.
\end{Definition}

We  conjecture that based on observation of Buchberger's algorithm's execution, an UNSAT core can be identified.
\begin{Proposition}
Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking S-poly computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.
\end{Proposition}
\begin{Example}
A SAT problem is described with 8 CNF clauses:

@@

% \begin{align*}
% &\bar{a}\lor\bar{b}\\
% &a\lor\bar{b}\\
% &\bar{a}\lor b\\
% &a\lor b\\
% &x\lor y\\
% &y\lor z\\
% &b\lor \neg y\\
% &a\lor x\lor \neg z
% \end{align*}
Using Boolean to polynomial mappings given in Table \ref{tab:booltof4}, we can transform them to a set of
polynomials $F$ over ring $\mathbb F_2[a,b,x,y,z]$:

\begin{minipage}[h]{0.4\textwidth}
\begin{align*}
&f_1:ab\\
&f_2:ab+a\\
&f_3:ab+b\\
&f_4:ab+a+b+1
\end{align*}
\end{minipage}
\begin{minipage}[h]{0.6\textwidth}
\begin{align*}
&f_5:xy+y+x+1\\
&f_6:yz+y+z+1\\
&f_7:by+y\\
&f_8:axz+az+xz+z
\end{align*}
\end{minipage}
% \begin{align*}
% &f_1:ab\\
% &f_2:ab+a\\
% &f_3:ab+b\\
% &f_4:ab+a+b+1\\
% &f_5:xy+y+x+1\\
% &f_6:yz+y+z+1\\
% &f_7:by+y\\
% &f_8:axz+az+xz+z
% \end{align*}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Fig.\ref{fig:UNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Fig.\ref{fig:UNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of clauses.
\end{enumerate}
\end{Example}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}
% We conclude our approach as follows: i) Keep track of all $Spoly(f_i,
% f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
% 0$; iii) Mark those $f_l$'s that have leading terms that cancel
% monomial terms in $Spoly(f_i, f_j)$; iv) Build a DAG recording all reductions including Spoly computations
% and multivariate divisions; v) Once $1\in J$ is
% detected, traverse the graph to identify all the marked polynomials that were
% used in obtaining this unit element. These polynomials constitute the
% core.

We conclude our approach as a conjecture algorithm (Algorithm \ref{alg:UNSAT}).

@@

\section{Formalize the Buchberger's Algorithm based UNSAT Core Identification}
\label{sec:core}

\textbf{Problem: }
Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
\emptyset$, or it is determined to be so by applying the Gr\"obner
basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
too. Borrowing the terminology from the Boolean SAT domain, we
call $F_c$ the infeasible core or the unsat core of $F$.

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to motivate that an unsat core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Def. \ref{def:gb}(ii)).

@@

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
$(f_2,f_5)$. It turns out that the Spoly-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Fig. \ref{fig:refute}. We call this graph a {\it refutation tree}.

\begin{figure}[hbt]
\centering
\includegraphics[width=\textwidth]{newfig/refutation_tree.eps}
\caption{Generating refutation trees to record unsat cores}
\label{fig:refute}
\end{figure}

In the figure, the nodes of the graph correspond to the polynomials
utilized in Buchberger's algorithm. The leaf nodes always correspond
to polynomials from the given generating set. An edge $e_{ij}$ from
node $i$ to node $j$ signifies that the polynomial at node $j$ results
from polynomial at node $i$. For example, consider the computation
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
following sequence of 1-step divisions:
$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}$. This is depicted
as the bottom subtree in the figure, terminating at polynomial
$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
division by $f_2$ resulted in the quotient $a$. The refutation tree of
Fig. \ref{fig:refute} shows further that
$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.

To identify an $F_c \subset F$, we start from the refutation node
``1'', and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an unsat core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Fig. \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an unsat core of $F$.

\section{Reducing the Size of the Infeasible Core $F_c$}
\label{sec:alg}
The core $F_c$ obtained from the aforementioned procedure may
contain redundant elements which could be discarded. For
example, consider the core $F_c=\{f_1,\dots, f_5\}$ generated in the
previous section. While $F_c$ is a smaller infeasible core of $F$, it
is not minimal. In fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
minimal core, where $F_{c4} \subset F_{c}$. Clearly, the polynomial
$f_3 \in F_c$ is a redundant element of the core and can be 
discarded. We will now describe techniques to further reduce the size
of the unsat core by identifying such redundant elements. 
%how the size of this core could be reduced further.  
For this purpose, we will have to perform a more 
systematic book-keeping of the data generated during the execution of
Buchberger's algorithm and the refutation tree.

@@

In Eqns. (\ref{eqn1}) and (\ref{data1}), $g_{ij}$ denotes the
remainder of the $S$-polynomial $Spoly(f_i,f_j)$ modulo the current system
of polynomials $f_1,\ldots,f_m$, and we denote by

$$h_{ij}:=\displaystyle\frac{LCM(lm(f_i),lm(f_j))}{lt(f_i)},
h_{ji}:= - \displaystyle\frac{LCM(lm(f_i),lm(f_j))}{lt(f_j)}$$ 
the coefficients of $f_i$, respectively $f_j$, in the $S$-polynomial
$Spoly(f_i,f_j)$. Furthermore, in Eqn. (\ref{data1}), $(c_{k1}, \dots
c_{kl})$ are the respective quotients of division by
polynomials $(f_{k1},\dots,f_{kl})$, generated during the $Spoly$ reduction.  
%polynomial coefficients of  that appear in the
%division process.

\begin{Example}
Revisiting Ex. \ref{ex:1}, and Fig. \ref{fig:refute}, the data
corresponding to $Spoly(f_1,f_2)$\\$\xrightarrow{F}_+  g_{12} = f_{10}$
reduction is obtained as the following sequence of computations:
$$f_{10}=g_{12}=f_1-acf_2-af_2-f_3-cf_2-f_2.$$ As the coefficient
field is $\mathbb{F}_2$ in this example, $-1 = +1$, so:
$$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$ is obtained.
The data is recorded according to Eqn. (\ref{data1}):

@@

Each element $g_i$ of $G$ is a polynomial combination of $\{f_1, \dots,
f_s\}$. Moreover, this matrix $M$ is constructed precisely using the
data that is recorded in the form of Eqn. (\ref{data1}). We now give a condition
when the matrix $M$ may identify some redundant elements.

@@

Corresponding to Example \ref{ex:1} and the refutation tree shown in
Fig. \ref{fig:refute}, we discover that the polynomial $f_3$ is used
only twice in the division process. In both occasions, the quotient of
the division is 1. From Fig. \ref{fig:refute}, it follows that:
\begin{equation}
1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + f_2)
\end{equation}

@@

{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
generators $\{f_1,\ldots,f_s\}$, then divide each of them by the
current basis $G=\{f_1,\dots,f_s,\dots,f_m\}$, which is the
intermediate result of Buchberger's algorithm.  
In this way, we obtain expressions of the following type:
\begin{equation}
\label{eqn:red}
g_{ij}= \underbrace{h_{ij}f_{i}+h_{ji}f_{j}}_{Spoly(f_i,f_j)}+\displaystyle\sum_{k=1}^m c_kf_k
\end{equation}
If the remainder $g_{ij}$ is non-zero, we denote it by
$f_{m+1}$ and add it to the current set of generators $G$. We
also record the data as in Eqn. (\ref{data1}): 
\begin{displaymath}
((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
This data forms a part of the refutation tree rooted at node $f_{m+1}$.

>>>>

\begin{table}[htb]
\centering
\begin{tabular}{|c|c|} 
\hline
 Boolean operator & Function over $\Fkk$ \\
\hline
\hline
 $a\land b$ & $a\cdot b$ \\
\hline
 $a\oplus b$ & $a+b$ \\
\hline
 $\bar{a}$ & $1+a$ \\
\hline
 $a\lor b$ & $a+b+a\cdot b$\\
\hline
\end{tabular}
\caption{Mapping Boolean operators to functions over $\Fkk$}
\label{tab:booltof4}  
\end{table}

@@

\subsection{Exploiting UNSAT cores for abstraction refinement}
Finding small UNSAT cores attracts interest for decades because of the need for all kinds of applications.
In solving MaxSAT problems, small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
It is applied to solving Boolean algebra problems, such as Boolean function decomposition \cite{lee2008bi}. 
Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.

UNSAT cores can find wide range of applications in circuit verification as well. Many abstraction refinement techniques require 
information mining from UNSAT proofs of intermediate abstractions. 
Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain how an UNSAT proof is
utilized in such techniques.

Bounded model checking (BMC) is a model checking technique which set upper bound to the length of all paths. 
It can be solved by solving a SAT problem. Given a model $M$, property $p$ and bound $k$, 
The $k$-BMC unrolls $M$ by $k$ clock-cycles and generates Boolean formula $F$ including violation check of $p$.
Then $F$ is fed to a SAT solver, if the SAT solver returns SAT, then $p$ is violated in 
some paths shorter than $k$. Otherwise, UNSAT indicates $p$ is not violated in all paths 
shorter than $k$. Algorithm \ref{alg:absrefine} makes an improvement on BMC using 
abstraction refinement.

@@

\begin{figure}[hbt]
\centering{
\includegraphics[width=5.5in]{newfig/refine.eps}
\caption{Abstraction by reducing latches}
\label{fig:refine}}
\end{figure}
Assume that we are given a sequential circuit with $n$ latches as shown in Figure \ref{fig:refine}(a). 
This circuit can be modeled as a Mealy machine $M$ and the states $s$ can be explicitly encoded by bit-level latch variables 
$l_1,\dots,l_n$. Algorithm \ref{alg:absrefine} describes an approach to check if machine $M$ violates property $p$.
This algorithm relies on $k$-BMC technique, which works on the basis of CNF-SAT solving.
The $k$-BMC represents the initial states $I$, the transition relation $T$ and property $p$ as CNF formulas.

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fixpoint detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf over-approximation}
of $M$, such that 
$$(M'\implies p) \implies (M\implies p)$$

If we find a violation on abstracted machine, then this
abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
when the size of datapath (number of latches in Figure \ref{fig:refine}) is very large. \textbf{Here we propose an altogether new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation:}

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While the complexity of computing a GB over finite fields is exponential in the number of variables,
the GB  computation is observed to be more efficient for UNSAT problems.} 
The reason is as follows:
\begin{Theorem}
{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high space and
time complexity and is usually not practical. One reason is that the size of the GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is $\{1\}$, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

Moreover, conventional techniques are limited to bit-level variables (literals). Algebraic geometry methods
allow the use of word-level variables, which provides a strong potential for all applications which can be
modeled as polynomials in finite fields extensions.

\section{UNSAT Cores of Polynomial Ideals}
\label{sec:core}
In this section we provide a solution to the problem that how to use 
GB to extract the UNSAT core.

\textbf{Problem statement: }
Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
\emptyset$, or it is determined to be so by applying the Gr\"obner
basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
too. Borrowing the terminology from the Boolean SAT domain, we
call $F_c$ the infeasible core or the unsat core of $F$.

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to motivate that an unsat core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Definition \ref{def:gb}(ii)).

\subsection{An Example Motivating our Approach}

Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking S-poly computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.

\begin{Example}
A SAT problem is described with 8 CNF clauses:

@@

\vspace{0.1in}
Using Boolean to polynomial mappings given in Table \ref{tab:booltof4}, we can transform them to a set of
polynomials $F=\{f_1,\dots,f_8\}$ over ring $\mathbb F_2[a,b,x,y,z]$:

\begin{minipage}[h]{0.4\textwidth}
\begin{align*}
&f_1:ab\\
&f_2:ab+a\\
&f_3:ab+b\\
&f_4:ab+a+b+1
\end{align*}
\end{minipage}
\begin{minipage}[h]{0.6\textwidth}
\begin{align*}
&f_5:xy+y+x+1\\
&f_6:yz+y+z+1\\
&f_7:by+y\\
&f_8:axz+az+xz+z
\end{align*}
\end{minipage}

\vspace{0.1in}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}[{1)}]
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:UNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:UNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of polynomials.
\end{enumerate}
\end{Example}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}
% We conclude our approach as follows: i) Keep track of all $Spoly(f_i,
% f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
% 0$; iii) Mark those $f_l$'s that have leading terms that cancel
% monomial terms in $Spoly(f_i, f_j)$; iv) Build a DAG recording all reductions including Spoly computations
% and multivariate divisions; v) Once $1\in J$ is
% detected, traverse the graph to identify all the marked polynomials that were
% used in obtaining this unit element. These polynomials constitute the
% core.

We conclude our approach as a conjecture Algorithm \ref{alg:UNSAT}.

@@

@@

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
$(f_2,f_5)$. It turns out that the Spoly-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Figure \ref{fig:refute}. We call this graph a {\it refutation tree}.

\begin{figure}[hbt]
\centering
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}
\caption{Generating refutation trees to record unsat cores}
\label{fig:refute}
\end{figure}

In the figure, the nodes of the graph correspond to the polynomials
utilized in Buchberger's algorithm. The leaf nodes always correspond
to polynomials from the given generating set. An edge $e_{ij}$ from
node $i$ to node $j$ signifies that the polynomial at node $j$ results
from polynomial at node $i$. For example, consider the computation
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
following sequence of 1-step divisions:
$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}$. This is depicted
as the bottom subtree in the figure, terminating at polynomial
$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
division by $f_2$ resulted in the quotient $a$. The refutation tree of
Figure  \ref{fig:refute} shows further that
$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.

To identify an $F_c \subset F$, we start from the refutation node
``1'', and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an unsat core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an unsat core of $F$.

\section{Reducing the Size of the Infeasible Core $F_c$}
\label{sec:alg}
The core $F_c$ obtained from the aforementioned procedure may
contain redundant elements which could be discarded. For
example, consider the core $F_c=\{f_1,\dots, f_5\}$ generated in the
previous section. While $F_c$ is a smaller infeasible core of $F$, it
is not minimal. In fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
minimal core, where $F_{c4} \subset F_{c}$. Clearly, the polynomial
$f_3 \in F_c$ is a redundant element of the core and can be 
discarded. We will now describe techniques to further reduce the size
of the unsat core by identifying such redundant elements. 
%how the size of this core could be reduced further.  
For this purpose, we perform a more 
systematic book-keeping of the data generated during the execution of
Buchberger's algorithm and the refutation tree.

@@

In Equation \ref{eqn1} and Equation \ref{data1}, $g_{ij}$ denotes the
remainder of the $S$-polynomial $Spoly(f_i,f_j)$ modulo the current system
of polynomials $f_1,\ldots,f_m$, and we denote by

$$h_{ij}:=\displaystyle\frac{LCM(lm(f_i),lm(f_j))}{lt(f_i)},
h_{ji}:= - \displaystyle\frac{LCM(lm(f_i),lm(f_j))}{lt(f_j)}$$ 
the coefficients of $f_i$, respectively $f_j$, in the $S$-polynomial
$Spoly(f_i,f_j)$. Furthermore, in Equation \ref{data1}, $(c_{k1}, \dots
c_{kl})$ are the respective quotients of division by
polynomials $(f_{k1},\dots,f_{kl})$, generated during the $Spoly$ reduction.  
%polynomial coefficients of  that appear in the
%division process.

\begin{Example}
Revisiting Example \ref{ex:1}, and Figure \ref{fig:refute}, the data
corresponding to $Spoly(f_1,f_2)$\\$\xrightarrow{F}_+  g_{12} = f_{10}$
reduction is obtained as the following sequence of computations:
$$f_{10}=g_{12}=f_1-acf_2-af_2-f_3-cf_2-f_2.$$ As the coefficient
field is $\mathbb{F}_2$ in this example, $-1 = +1$, so:
$$f_{10}=g_{12}=f_1+acf_2+af_2+f_3+cf_2+f_2$$ is obtained.
The data is recorded according to Equation \ref{data1}:

@@

Each element $g_i$ of $G$ is a polynomial combination of $\{f_1, \dots,
f_s\}$. Moreover, this matrix $M$ is constructed precisely using the
data that is recorded in the form of Equation \ref{data1}. We now give a condition
when the matrix $M$ may identify some redundant elements.

@@

Corresponding to Example \ref{ex:1} and the refutation tree shown in
Figure \ref{fig:refute}, we discover that the polynomial $f_3$ is used
only twice in the division process. In both occasions, the quotient of
the division is 1. From Figure \ref{fig:refute}, it follows that:
\begin{equation}
1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + f_2)
\end{equation}

@@

{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
generators $\{f_1,\ldots,f_s\}$, then divide each of them by the
current basis $G=\{f_1,\dots,f_s,\dots,f_m\}$, which is the
intermediate result of Buchberger's algorithm.  
In this way, we obtain expressions of the following type:
\begin{equation}
\label{eqn:red}
g_{ij}= \underbrace{h_{ij}f_{i}+h_{ji}f_{j}}_{Spoly(f_i,f_j)}+\displaystyle\sum_{k=1}^m c_kf_k
\end{equation}
If the remainder $g_{ij}$ is non-zero, we denote it by
$f_{m+1}$ and add it to the current set of generators $G$. We
also record the data as in Equation \ref{data1}: 
\begin{displaymath}
((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
This data forms a part of the refutation tree rooted at node $f_{m+1}$.

{\bf Recording the coefficients:} In Equation \ref{eqn:red} we obtain a
vector of polynomial coefficients $c_k$ where $k>s$. These
coefficients are associated with new elements (remainders) in the
Gr\"obner basis, that are not a part of the unsat core. 
%which cannot benefit the UNSAT core extraction. 
Since each polynomial $f_k$, ($k>s$) is generated by
$\{f_1,\dots,f_s\}$, we can re-express $f_k$ in terms of $\{f_1,\dots,
f_s\}$. Thus, each $f_k, k>s$ can be written as $f_k = d_1f_1 + \dots
+ d_sf_s$. This process adds a new row $(d_1,\dots,d_s)$ to the
coefficient matrix $M$.
